---
title: "HW3"
author: "ManBuXiaoCun"
date: '2022-05-03'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(glue)
library(tidyverse)
library(lubridate)
library(stringr)
```

第三周作业
说明：自定义函数至少做一组测试

1. 自定义函数根据身高、体重，计算BMI以及按国内标准的分类。
注：使用多返回值，附
BMI = kg/m2
If your BMI is less than 18.5, it falls within the **underweight** range.
If your BMI is 18.5 to 24.9, it falls within the **normal** or Healthy Weight range.
If your BMI is 25.0 to 29.9, it falls within the **overweight** range.
If your BMI is 30.0 or higher, it falls within the **obese** range.
```{r}
bmi <- function(height = 180, weight = 80){
  bmi_index <- weight/(height/100)^2
  bmi_index <- round(bmi_index, digits = 2)
  if (bmi_index < 18.5 ){
   glue(" BMI IS {bmi_index}, underweight")
    
  } else if (bmi_index < 24.9 ){
    glue(" BMI IS {bmi_index}, normal")
    
  } else if(bmi_index < 29.9 ){
   glue(" BMI IS {bmi_index}, overweight")
    
  } else {
    glue(" BMI IS {bmi_index}, obese")
  }
  
  
}
```


2. 根据课表数据框，
(1) 用for循环计算每天的学时数；
(2) 自定义函数，根据课表数据框计算每天学时数。

3. 自定义函数根据身份证号，计算两种年龄、性别。
   注：使用默认参数和多返回值。
```{r}
# id has to be in quote when using the function
# unknown how to handle the custmize err message to remind function user 
  id_age_gender <- function(id="51162119920320211X"){
  
  # age
  birthday <- str_sub(id, start = 7,end = 14)
  birthday <- ymd(birthday) # change into year month day
  age <- year(today())-year(birthday)
  # gender
  gender_num <- str_sub(id, start = 17,end = 17) # extract gender identity num
  gender_num <- as.numeric(gender_num) # convert character to number
  gender <- if_else(gender_num %% 2 == 0,"Female","Male") # convert num to male and female
  
  # output part
  glue("This year is {year(today())},You are {age} years old, Gender: {gender}")
  
}
```


4. 将第3题中的函数做向量化改进：
(1) 修改函数体，使其接受向量化输入
(2) 借助map系列实现
并用随机生成的3个身份证号测试：
"11022119920110474X"
"230702198105052397" 
"142625199601063388"
```{r}
# not solved yet
id_nums <- list("11022119920110474X",
"230702198105052397", 
"142625199601063388")
map(id_nums,id_age_gender())
```


5. 自定义函数实现经典猜数游戏：随机生成一个0~300的整数，玩家共有5次机会，每次猜数都会返回提示：猜大了还是猜小了，并提示玩家还胜几次机会。
提示：借助函数readline()接受用户输入。
```{r}
hidden_number <- 3 #sample(0:6,size=1)
guess_num <- readline(prompt = "please input a number ")
while (guess_num!=hidden_number) {
  
}
if(guess_num > hidden_number)
  {readline(prompt = "please input a smaller number ")
}else if(guess_num > hidden_number){
  readline(prompt = "please input a larger number ")  
}
else{print("You win")}

```
EXAMPLES from website
https://setscholars.net/r-examples-for-beginners-guess-a-random-number-game-in-r/
```{r}
readinteger <- function()
{ 
  n <- readline(prompt="Enter an integer: ")
  if(!grepl("^[0-9]+$",n))
  {
    return(readinteger())
  }
  return(as.integer(n))
}

# real program start here

num <- sample(1:300,size = 1)
guess <- -1

cat("Guess a number between 0 and 100.n")

while(guess != num)
{ 
  guess <- readinteger()
  if (guess == num)
  {
    cat("Congratulations,", num, "is right.n")
  }
  else if (guess < num)
  {
    cat("It's bigger!n")
  }
  else if(guess > num)
  {
    cat("It's smaller!n")
  }
}
```


